---
title: "Lab 03 - Nobel laureates"
author: "Marcus Minko"
date: "01-25-2022"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1
```{r data-glimpse}
glimpse(nobel)
```
There is one row for each laureate. 26 variables and 935 observations.

### Exercise 2
```{r nobel-living}
nobel_living <- nobel %>% 
  filter(
    gender != "org", 
    country != "NA", 
    is.na(died_date)
    )

glimpse(nobel_living)
```


```{r mutate-usa}
nobel_living <- nobel_living %>%
  mutate(
    country_us = if_else(country == "USA", "USA", "Other")
  )

# nobel living science dataframe

nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))

```

### Exercise 3
```{r data-viz}
ggplot(data = nobel_living_science, 
      mapping = aes(x = country_us,
                    color = country_us, fill = country_us)) +
    geom_bar() +
    coord_flip() +
    facet_wrap(~ category) +
  labs(title = "USA vs. non-USA Laureates by Discipline", 
       x = "Country",
       y = "# of nobel winners") 
  
```
Yes, there are more winners from USA than other countries.
...

### Exercise 4

```{r mutate-USA-born}
nobel_living_science <- nobel_living_science %>%
  mutate(
    born_country_us = if_else(born_country == "USA", "USA", "Other")
  ) 
nobel_living_science %>%
  filter(born_country_us == "USA")
```

105 of the winners are USA born.
...

### Exercise 5

```{r plot-country-born}
ggplot(data = nobel_living_science, 
      mapping = aes(x = country_us,
                    fill = born_country_us)) +
    geom_bar() +
    coord_flip() +
    facet_wrap(~ category) +
  labs(title = "USA vs. non-USA Laureates by Disicipline", 
       x = "Living in US or other",
       y = "# of nobel winners") 
```

A substantial majority of the USA nobel winners were born in the USA and thus did not themselves immigrate.
...

### Exercise 6

```{r frequency-other-usa}
nobel_living_science %>%
  filter(country == "USA" & born_country_us == "Other") %>%
  count(born_country) %>%
  arrange(desc(n))
```
UK and Germany are tied for the most common USA nobel winners born elsewhere.
...
